---
title: "CACTI Study CVD Measures"
author: "Laura Pyle, PhD"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(readxl)
library(magrittr)
library(purrr)
library(readr)
library(tableone)
library(Hmisc)
library(labelled)

# read in raw data
cat <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\CACTI\\CACTI oral microbiome\\Blinded CSOC report\\Data\\Blinded category.csv")
pwa <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\CACTI\\CACTI oral microbiome\\Blinded CSOC report\\Data\\PWA.csv")
pwv <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\CACTI\\CACTI oral microbiome\\Blinded CSOC report\\Data\\PWV.csv")
cIMT <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\CACTI\\CACTI oral microbiome\\Blinded CSOC report\\Data\\cIMT.csv")
BrachD <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\CACTI\\CACTI oral microbiome\\Blinded CSOC report\\Data\\BrachD.csv")

# clean up extra rows
cat <- cat[!is.na(cat$StudyID),] # N=607
pwa <- pwa[!is.na(pwa$StudyID),]
pwv <- pwv[!is.na(pwv$StudyID),]
cIMT <- cIMT[!is.na(cIMT$StudyID),]
BrachD <- BrachD[!is.na(BrachD$StudyID),]

# for cIMT, take largest of left and right mean, unless only one side
for (i in 1:nrow(cIMT)) {
  cIMT$cIMT[i] <- ifelse(is.na(cIMT$CIMT_LeftMean[i]) & is.na(cIMT$CIMT_RightMean[i]),NA,max(cIMT$CIMT_LeftMean[i],cIMT$CIMT_RightMean[i],na.rm = TRUE))
}

# for PWV, PWA, BD, average per person
pwv_mean <- aggregate(pwv[,4:5],by=list(pwv$StudyID),mean,na.rm=TRUE)
colnames(pwv_mean) <- c("StudyID","MeanHeartRate","MeanPulseWaveVelocity")
pwv_mean <- pwv_mean[!is.na(pwv_mean$MeanHeartRate),]

pwa_mean <- aggregate(pwa[,4:6],by=list(pwa$StudyID),mean,na.rm=TRUE)
colnames(pwa_mean) <- c("StudyID","C_Ap","C_Ph","C_Agph")
pwa_mean <- pwa_mean[!is.na(pwa_mean$C_Ph),]

bd_mean <- aggregate(BrachD[,4],by=list(BrachD$StudyID),mean,na.rm=TRUE)
colnames(bd_mean) <- c("StudyID","BA_Distensibilty")
bd_mean <- bd_mean[!is.na(bd_mean$BA_Distensibilty),]

# merge data by StudyID
alldata <- merge(cat,pwa_mean,by="StudyID",all.x=TRUE,all.y=TRUE)
alldata <- merge(alldata,pwv_mean,by="StudyID",all.x=TRUE,all.y=TRUE)
alldata <- merge(alldata,bd_mean,by="StudyID",all.x=TRUE,all.y=TRUE)
alldata <- merge(alldata,cIMT,by="StudyID",all.x=TRUE,all.y=TRUE)
alldata[alldata$StudyID=="",] <- NA
alldata <- alldata[!is.na(alldata$StudyID),]

# currently there is one ppt with Group.Status="N" - take them out for now
alldata <- alldata[alldata$Group.Status=="A" | alldata$Group.Status=="B",]
alldata$Group.Status <- factor(alldata$Group.Status)

# some people have 0 values for BD, set to missing
alldata$BA_Distensibilty[alldata$BA_Distensibilty==0] <- NA

# get rid of people missing all data 
alldata <- alldata[!is.na(alldata$C_Ap) | !is.na(alldata$C_Ph) | !is.na(alldata$C_Agph) | !is.na(alldata$MeanHeartRate) |
                     !is.na(alldata$MeanPulseWaveVelocity) | !is.na(alldata$BA_Distensibilty) | !is.na(alldata$cIMT),]

# count number of each measure
n_C_Ap <- nrow(alldata[!is.na(alldata$C_Ap),])
n_C_Ph <- nrow(alldata[!is.na(alldata$C_Ph),])
n_C_Agph <- nrow(alldata[!is.na(alldata$C_Agph),])
n_HR <- nrow(alldata[!is.na(alldata$MeanHeartRate),])
n_PWV <- nrow(alldata[!is.na(alldata$MeanPulseWaveVelocity),])
n_BD <- nrow(alldata[!is.na(alldata$BA_Distensibilty),])
n_cIMT <- nrow(alldata[!is.na(alldata$cIMT),])

# labels
var_label(alldata$cIMT)="cIMT (mm)"
var_label(alldata$C_Ap)="Augmented Pressure (mmHg)"
var_label(alldata$C_Agph)="Augmentation Index (%)"
var_label(alldata$BA_Distensibilty)="Brachial Distensibility (%/mmHg)"
var_label(alldata$MeanPulseWaveVelocity)="Pulse Wave Velocity (m/s)"

myvars <- c("C_Ap","C_Agph","MeanPulseWaveVelocity","BA_Distensibilty","cIMT")
tab <- CreateTableOne(vars=myvars, strata="Group.Status",data=alldata, test=FALSE)
tab <- print(tab,varLabels=TRUE)

```

# Background

The CACTI Study Clinical Study Oversight Committee (CSOC) is concerned that there may not be enough subclinical cardiovascular disease (CVD) in the non-diabetic group to answer the research question.  The CSOC has requested a closed report with means and standard deviations of all CVD measures, stratified by study group, as part of the interim report.  This table should be developed and provided by a blinded statistician.

# Methods

The CVD outcomes data were provided to the blinded statistician, along with blinded group memberships.  Participants with at least one CVD measure were included in this report.  For cIMT, the maximum of the left and right sides was used in analysis.  For pulse wave velocity, Augmentation Index, augmented pressure, and brachial distensitiblity, the mean of a participant's multiple measures was used in analysis.

Participants were stratified by group and descriptive statistics were calculated.

# Results

Table 1 provides descriptive statistics for the CVD measures, by blinded study group.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kableone(tab,caption="Table 1.  Descriptive statistics for CVD measures, by blinded study group.")
```
