---
title: "CACTI Study CVD Measures"
author: "Laura Pyle, PhD"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(readxl)
library(magrittr)
library(purrr)
library(readr)
library(tableone)
library(Hmisc)
library(labelled)

# read in Janet's combined file
alldata <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\CACTI\\CACTI oral microbiome\\Blinded CSOC report #2\\Data\\QOMBE_Blinded32320.csv")

# read in augmentation index file
ai <- read.csv("S:\\Shared Projects\\Laura\\BDC\\Projects\\CACTI\\CACTI oral microbiome\\Blinded CSOC report #2\\Data\\PWA_Query.csv")
ai$VisitDate <- ai$StudyDateTime
ai <- ai[,c("StudyID","StudyDateTime","C_Ai")]

# clean up extra rows
alldata <- alldata[!is.na(alldata$StudyID),] 

# for cIMT, take largest of left and right mean, unless only one side
for (i in 1:nrow(alldata)) {
  alldata$cIMT[i] <- ifelse(is.na(alldata$CIMT_LeftMean[i]) & is.na(alldata$CIMT_RightMean[i]),NA,max(alldata$CIMT_LeftMean[i],alldata$CIMT_RightMean[i],na.rm = TRUE))
}

# some people have 0 values for BD, set to missing
alldata$BrachD[alldata$BrachD==0] <- NA

# count number of each measure
n_Ap <- nrow(alldata[!is.na(alldata$Ap),])
n_Ph <- nrow(alldata[!is.na(alldata$Ph),])
n_Agph <- nrow(alldata[!is.na(alldata$Agph),])
n_HR <- nrow(alldata[!is.na(alldata$HR),])
n_PWV <- nrow(alldata[!is.na(alldata$PWV),])
n_BD <- nrow(alldata[!is.na(alldata$BrachD),])
n_cIMT <- nrow(alldata[!is.na(alldata$cIMT),])

# now classify as normal/abnormal
alldata$cIMT_abnormal <- ifelse(is.na(alldata$cIMT),NA,ifelse(alldata$cIMT>1,1,0))
alldata$PWV_abnormal <- ifelse(is.na(alldata$PWV),NA,
                               ifelse(alldata$AgeV1>=30 & alldata$AgeV1<40 & alldata$PWV>6.47,1,
                                      ifelse(alldata$AgeV1>=40 & alldata$AgeV1<50 & alldata$PWV>7.03,1,
                                             ifelse(alldata$AgeV1>=50 & alldata$AgeV1<60 & alldata$PWV>8.33,1,
                                                    ifelse(alldata$AgeV1>=60 & alldata$AgeV1<70 & alldata$PWV>8.68,1,
                                                    ifelse(alldata$AgeV1>=70 & alldata$AgeV1<80 & alldata$PWV>9.76,1,0))))))
alldata$BrachD_abnormal <- ifelse(is.na(alldata$BrachD) | !(alldata$RaceVal %in% c("White","Black or African American")),NA,
                              ifelse(alldata$RaceVal=="White" & alldata$Gender=="M" & alldata$BrachD <6.38,1,
                                ifelse(alldata$RaceVal=="White" & alldata$Gender=="F" & alldata$BrachD <6.7,1,
                                  ifelse(alldata$RaceVal=="Black or African American" & alldata$Gender=="M" & alldata$BrachD <6,1,
                                         ifelse(alldata$RaceVal=="Black or African American" & alldata$Gender=="F" & alldata$BrachD<6.25,1,0)))))

# labels
var_label(alldata$cIMT)="cIMT (mm)"
var_label(alldata$Ap)="Augmented Pressure (mmHg)"
var_label(alldata$Agph)="Augmentation Index (%)"
var_label(alldata$BrachD)="Brachial Distensibility (%/mmHg)"
var_label(alldata$PWV)="Pulse Wave Velocity (m/s)"

myvars <- c("Ap","Agph","PWV","BrachD","cIMT")
tab <- CreateTableOne(vars=myvars, strata="AStatus",data=alldata, test=FALSE)
tab <- print(tab,varLabels=TRUE)

```

# Background

The CACTI Study Clinical Study Oversight Committee (CSOC) is concerned that there may not be enough subclinical cardiovascular disease (CVD) in the non-diabetic group to answer the research question.  The CSOC has requested a closed report with means and standard deviations of all CVD measures, stratified by study group, as part of the interim report.  This table should be developed and provided by a blinded statistician.

# Methods

The CVD outcomes data were provided to the blinded statistician, along with blinded group memberships.  Participants with at least one CVD measure were included in this report.  For cIMT, the maximum of the left and right sides was used in analysis.  For pulse wave velocity, Augmentation Index, augmented pressure, and brachial distensitiblity, the mean of a participant's multiple measures was used in analysis.

Participants were stratified by group and descriptive statistics were calculated.

# Results

Table 1 provides descriptive statistics for the CVD measures, by blinded study group.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kableone(tab,caption="Table 1.  Descriptive statistics for CVD measures, by blinded study group.")
```
